---
title: "HW 3"
author: "Sachiko Lamen and Chase Tarr"
date: "5/8/2022"
output:
  html_document:
    code_folding: hide
  pdf_document: default
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(rootSolve)
library(thematic)
library(scales)
library(equatiomatic)
```

```{r}
# Read in data and clean 
data <- read_csv(here("data",'HW3_data.csv')) %>%
  select(-1) %>% 
  clean_names()
# Make income a variable, not a column
data_long <- data %>% 
  pivot_longer(cols = c(q_low_kwh, q_high_kwh),
               names_to = 'income_level',
               values_to = 'kwh') %>% 
  mutate(income_level = case_when(income_level == 'q_low_kwh' ~ 'low',
                   income_level == 'q_high_kwh' ~ 'high'))
```

## 1. One kWh of electricity emits 0.85 lbs of $CO_2$. Assuming the interim SCC correctly reflects the total social cost of one metric ton of $CO2$, what is the marginal externality cost per kWh of electricity?

```{r}
# Calculate MEC of carbon, assuming SCC = $51/metric ton, 1 kWh = 0.85lbs CO2
MEC <- (0.85*51)/2204.62
```

MEC is $`r round(MEC, 2)` per kWh

## 2. What is the aggregate monthly demand curve for electricity? What is the supply curve for electricity? What is the "benefit" to consumers under the status quo? What is the "benefit" to producers under the status quo? What is the environmental cost under the status quo?

```{r, include=FALSE}
# Create linear models for low demand
demand_low <- lm(price_cents ~ kwh, income_level == "low", data = data_long)
low <- extract_eq(model = demand_low, use_coefs = TRUE, coef_digits = 5)
# Create linear models for high demand
demand_high <- lm(price_cents ~ kwh, income_level == "high", data = data_long)
high <- extract_eq(model = demand_high, use_coefs = TRUE, coef_digits = 5)
# need to rearrange the parameter to get Q(P)! 
# Qgg = Qlow(P) + Qhigh(P) 
# Importantly, since they-intercepts are different, we know that Qagg(P) will have a kink. I include an ifelse() statement to take
# care of the kink.
# define a function for demand 
demand <- function(p, model){
  q <- (p - model$coefficients[[1]])/model$coefficients[[2]]
  q <- ifelse(q < 0, 0, q)
  return(q)
}
# Find aggregate demand for each P value
demand_agg <- function(p){
  q <- demand(p, demand_low) + demand(p, demand_high)
  return(q)
}
# Set price vector (0, 30)
price = seq(0, 30, length.out = 100)
# Define aggregate quantity
kwh <- map(price, demand_agg) %>% unlist()
# Make data frame of  price and quantity values for aggregate curve
data_agg <- tibble(kwh = kwh, price = price)
# Create linear model for aggregate demand
agg_d_lm <- lm(price ~ kwh, data = data_agg)
agg <- extract_eq(model = agg_d_lm, use_coefs = TRUE, coef_digits = 5)
agg_intercept <- agg_d_lm$coefficients[[1]]
agg_slope <- agg_d_lm$coefficients[[2]]
# Supply
supply <- demand_agg(10)
# 536719.5 at $0.10
supply_slope <- 10/supply
#0.0000186
supply_function <- function(p){
  q <- p/supply_slope
  return(q)
}
# supply curve is going to pass through agg demand curve at $0.10/kwh, the slope of supply will be 1.83x10^-5 x q
P <- 10
# Creating MPC function
MPC <- function(q){
  L <- (q*MPC-supply_slope)
  return(L)
}
```


```{r, warning = FALSE}
data %>%
  mutate(d_low = predict(demand_low)) %>%
  mutate(d_high = predict(demand_high)) %>%
  ggplot() +
  geom_point(aes(q_low_kwh, price_cents), color = "green") + 
  geom_line(aes(q_low_kwh, d_low), color = "green") +
  geom_point(aes(q_high_kwh, price_cents), color = "blue") +
  geom_line(aes(q_high_kwh, d_high), color = "blue") +
  geom_line(data = data_agg, aes(kwh, price), color = "red") +
  geom_hline(yintercept = 10) +
  geom_abline(color = "darkorchid4",
              intercept = 0,
              slope = supply_slope)
  theme_minimal() +
  labs( x = "Electricity (kWh)", y = "Price (cents/kWh)")
```

**Figure 1.** Low demand is represented by the green line. High demand is represented by the blue line. Aggregate demand is represented by the red line. Supply curve is represented by purple (aka...darkorchid4 ;))

**Low demand:** `r low`

**High demand:** `r high`

**Aggregate demand: ** `r agg`

**Supply** $P = 10$ *(note that price is in cents/kWh)*

```{r}
# Write CS function
CS <- function(price_cents, model){
  q <- demand(price_cents, model)
  cs <- 0.5*(model$coefficients[[1]] - price_cents)*q
  return(cs)
}
CS_agg <- function(price_cents){
  cs <- CS(price_cents, demand_low) + CS(price_cents, demand_high)
  return(cs)
}
# Find Q and P at intersection of aggregate demand and supply curves 
Q0 <- agg_intercept/(10 + agg_intercept)
P0 <- 10
# Calculate "benefit" to consumers under status quo
CS_sq_ct <- 0.5*supply*(0.305 - 0.1)
CS_sq_ct
# consumer benefit : $55,013.75
# Calculate "benefit" to producers under status quo
PS_sq_ct <- 0.5*supply*.1
PS_sq_ct
# producer benefit: $26,106
# Calculate environmental cost under status quo
env_cost_sq_ct <- MEC*supply
env_cost_sq_ct
# environmental cost: $10,265
```
**Q at status quo:** `r Q0`

**P at status quo:** `r P0`

**Benefit to Consumers (status quo):** `r CS_sq_ct`

**Benefit to Producers (status quo):** `r PS_sq_ct`

**Environmental Cost (status quo):** `r env_cost_sq_ct`

3. How is the current consumer benefit divided between "high" and "low" income consumers?

```{r}
CS_low <- CS(10, demand_low) 
CS_high <- CS(10, demand_high)
```
At status quo, **high income** consumers benefit `r CS_high` while **low income** consumers benefit only `r CS_low`. Overall, the **difference in benefit** is `r CS_high - CS_low`

4. Derive the optimal electricity tax (in cents/kWh) using the interim SCC. Noting that recent research has show the poor face a disproportionate share of the impacts from climate change, assume that the climate externality is born entirely by the "low" income group. What would be the effects of this tax on:
```{r}
# know that P* = 10
optimal_tax <- MEC
# taxing the consumer (household) [P = mQ + b]
```

```{r}
#elec_prod <- MPC()
tax <- 1.97
demand_tax <- function(p, model){
  q <- (p - (model$coefficients[[1]]-tax))/model$coefficients[[2]]
  q <- ifelse(q < 0, 0, q)
  return(q)
}
demand_agg_tax <- function(p){
  q <- demand_tax(p, demand_low) + demand_tax(p, demand_high)
  return(q)
}
# calculate when the aggregate demand intersects the supply curve
uniroot(function(p)
  demand_agg_tax(p) - supply_function(p),
  interval = c(0,20))

p_new_a <- 9.32
q_new_a <- demand_agg_tax(p_new_a)


#new equilibrium price and quantities
demand_change_tax <- supply-q_new_a
price_change_tax <- 10 - p_new_a
```

```{r}
demand_high_tax <- demand_tax(p_new_a, demand_high)
cs_high_tax <- (0.5*(demand_high$coefficients[1]-tax-p_new_a)*demand_high_tax/100)
high_welfare_tax <- (0.5*(0.3161-0.0197-0.0932)*demand_high_tax)-(demand_high_tax * 0.0197)
overall_welfare_high_cons <- cs_high_tax - high_welfare_tax
```

```{r}
#find change from original high income consumer surplus - welfare high tax
original_welfare_low <- CS_low-env_cost_sq_ct
demand_tax_low <- demand_tax(p_new_a, demand_low)
env_cost_tax <- MEC*q_new_a
cs_low_tax <- (0.5*(demand_low$coefficients[1]-tax-p_new_a)*demand_tax_low)/100
welfare_tax_low <- cs_low_tax-(demand_tax_low*0.0197)-env_cost_tax
welfare_low_change<- original_welfare_low-welfare_tax_low
```

```{r}
ps_tax <- 0.5*q_new_a*0.0932
ps_change <- PS_sq_ct-ps_tax
```

```{r}
env_dam <- env_cost_sq_ct-env_cost_tax
```

```{r}
tax_rev <- MEC*q_new_a
```

a) The amount of electricity produced and consumed:
The amount of electricity produced/consumed will decrease by `r round(demand_change_tax, 0)`kWh. Total consumption after change will be `r round(q_new_a, 0)`

b) The price of electricity:
Price of electricity is 9.32 /kWh. Price will decrease by $`r round(price_change_tax, 2)`. 

c) The overall welfare of "high" income consumers:
High income welfare will decrease by $`r round(overall_welfare_high_cons, 0)`. 

d) Overall welfare of "low" income consumers:
Low-income welfare will decrease by $`r round(welfare_low_change, 0)`. 

e) Power Suppliers (i.e., electricity producers):
Producer surplus will decrease by $ `r round(ps_change, 0)`.

f) Total environmental damage:
Cost of environmental damage will decrease$`r round(env_dam, 0)`.

g) Total tax revenue generated:
Amount of tax revenue generated: $`r round(tax_rev, 0)`. 

5. Now, assume that all revenue from the electricity tax will be redistributed to the consumers in proportion to their pre-tax consumption. For example, if 80% of the electricity was consumed by “high” income consumers, then they get 80% of the tax revenue. Additionally, consider the fact that current scientific evidence suggests the true SCC may be much higher than 51. For a range of SCC values (51, 75, 100, 125, and 150 per metric ton of CO2), calculate the effects of an SCC-based electricity tax on:

```{r}
value <- 2204.6/0.86
tax75 <- 75/(value)
tax100 <- 100/value
tax125 <- 125/value
tax150 <- 150/value
```

###### 51
```{r}
# vector of SCC to calculate welfare
SCC_vec <- c(2.89,3.86,4.82,5.78)

low_percent <- demand(10,demand_low)/supply
high_percent <- demand(10, demand_high)/supply

low_distr_51 <- low_percent*tax_rev
low_welfare_51 <- low_distr_51 + welfare_tax_low

high_distr_51 <- high_percent*tax_rev
high_welfare_51 <- high_distr_51 + welfare_tax_low
```

######75
```{r}
# create a new function to find p values of 75-150
scc_d <- function(p, model, scc){
  q <- (p-(model$coefficients[[1]]-SCC_vec[scc]))/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}
agg_d_75 <- function(p){
  q <- scc_d(p, demand_low, 1) + scc_d(p, demand_high, 1)
  return(q)
}
p_75 <- uniroot(function(p)
  agg_d_75(p) - supply_function(p),
  interval = c(0,20))
demand_75 <- agg_d_75(p_75$root)
rev_loss75 <- demand_75*.0289
low_d_75 <- scc_d(p_75$root,demand_low,1)
cs_low75 <- (0.5*low_d_75*(demand_low$coefficients[1]-tax75 - p_75$root))/100
welfare_low75 <- cs_low75 - (low_d_75*0.0289) - rev_loss75 + low_percent*rev_loss75
high_d75 <- scc_d(p_75$root,demand_high,1)
high_cs75 <- (0.5*high_d75*(demand_high$coefficients[1]-tax75-p_75$root))/100
high_welfare_75 <- high_cs75 - (high_d75*0.0289) + high_percent*rev_loss75
ps_75 <- (0.5*demand_75*(p_75$root/100))
```

######100
```{r}
agg_d_100 <- function(p){
  q <- scc_d(p, demand_low, 2) + scc_d(p, demand_high, 2)
  return(q)
}
p_100 <- uniroot(function(p)
  agg_d_100(p) - supply_function(p),
  interval = c(0,20))
demand_100 <- agg_d_100(p_100$root)
rev_loss100 <- demand_100*.0386
low_d_100 <- scc_d(p_100$root,demand_low,2)
cs_low100 <- (0.5*low_d_100*(demand_low$coefficients[1]-tax75 - p_100$root))/100
welfare_low100 <- cs_low100 - (low_d_100*0.0386) - rev_loss100 + low_percent*rev_loss100
high_d100 <- scc_d(p_100$root,demand_high,2)
high_cs100 <- (0.5*high_d100*(demand_high$coefficients[1]-tax100-p_100$root))/100
high_welfare_100 <- high_cs100 - (high_d100*0.0386) + high_percent*rev_loss100
ps_100 <- (0.5*demand_100*(p_100$root/100))
```

######125
```{r}
agg_d_125 <- function(p){
  q <- scc_d(p, demand_low, 3) + scc_d(p, demand_high, 3)
  return(q)
}
p_125 <- uniroot(function(p)
  agg_d_125(p) - supply_function(p),
  interval = c(0,20))
demand_125 <- agg_d_125(p_125$root)
rev_loss125 <- demand_125*.0482
low_d_125 <- scc_d(p_125$root,demand_low,3)
cs_low125 <- (0.5*low_d_125*(demand_low$coefficients[1]-tax125 - p_125$root))/100
welfare_low125 <- cs_low125 - (low_d_125*0.0482) - rev_loss125 + low_percent*rev_loss125
high_d125 <- scc_d(p_125$root,demand_high,3)
high_cs125 <- (0.5*high_d125*(demand_high$coefficients[1]-tax125-p_125$root))/100
high_welfare_125 <- high_cs125 - (high_d125*0.0482) + high_percent*rev_loss125
ps_125 <- (0.5*demand_125*(p_125$root/100))
```

######150
```{r}
agg_d_150 <- function(p){
  q <- scc_d(p, demand_low, 4) + scc_d(p, demand_high, 4)
  return(q)
}
p_150 <- uniroot(function(p)
  agg_d_150(p) - supply_function(p),
  interval = c(0,20))
demand_150 <- agg_d_150(p_150$root)
rev_loss150 <- demand_150*.0578
low_d_150 <- scc_d(p_150$root,demand_low,4)
cs_low150 <- (0.5*low_d_150*(demand_low$coefficients[1]-tax150 - p_150$root))/100
welfare_low150 <- cs_low150 - (low_d_150*0.0578) - rev_loss150 + low_percent*rev_loss150
high_d150 <- scc_d(p_150$root,demand_high,4)
high_cs150 <- (0.5*high_d150*(demand_high$coefficients[1]-tax150-p_150$root))/100
high_welfare_150 <- high_cs150 - (high_d150*0.0578) + high_percent*rev_loss150
ps_150 <- (0.5*demand_150*(p_150$root/100))
```

a) Overall welfare of  "high" income consumers:
- SCC 51 USD/ton: $`r round(high_welfare_51, 0)`. 
- SCC 75 USD/ton: $`r round(high_welfare_75, 0)`. 
- SCC 100 USD/ton: $`r round(high_welfare_100, 0)`.
- SCC 125 USD/ton: $`r round(high_welfare_125, 0)`.
- SCC 150 USD/ton: $`r round(high_welfare_150, 0)`.

b) Overall welfare of  "low" income consumers:

- SCC 51 USD/ton: $`r round(low_welfare_51, 0)`. 
- SCC 75 USD/ton: $`r round(welfare_low75, 0)`. 
- SCC 100 USD/ton: $`r round(welfare_low100, 0)`.
- SCC 125 USD/ton: $`r round(welfare_low125, 0)`.
- SCC 150 USD/ton: $`r round(welfare_low150, 0)`.

c) Electricity Producers:

- SCC 51 USD/ton: $`r round(ps_tax, 0)`. 
- SCC 75 USD/ton: $`r round(ps_75, 0)`. 
- SCC 100 USD/ton: $`r round(ps_100, 0)`.
- SCC 125 USD/ton: $`r round(ps_125, 0)`.
- SCC 150 USD/ton: $`r round(ps_150, 0)`.

6. Suppose the “high” income group has access to expensive home solar generation. This lowers the electricity demand curve for the “high” income group by half (vertically). Under this new demand:

```{r}
high_solar_d<- function(p){
  q <- (p - (demand_high$coefficients[[1]]-1.97)/2)/demand_high$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}
low_solar_d <- function(p){
  q <- (p - (demand_low$coefficients[[1]]-1.97))/demand_low$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}
agg_solar_d <- function(p){
  q <- high_solar_d(p) + low_solar_d(p)
  return(q)
}
solar_peepee <- uniroot(function(p)
  agg_solar_d(p) - supply_function(p),
        interval = c(0,20))
p_solar <- solar_peepee$root
solar_d <- agg_solar_d(p_solar)
solar_env_cost <- solar_d*0.0197
finaltax <- (.0197/(solar_env_cost/env_cost_tax))
```

(a) What is total electricity consumption?
`r round(solar_d, 0)` kWh.

(b) What is the total environmental externality?
$`r round(solar_env_cost, 0)`

(c) What value of the electricity tax makes the total environmental damage the same as the damage when solar panels are available to the high income group? 

$`r round(finaltax, 4)`.